---
title: "Info 201 FP Group 2"
author: "Zachary Wu, Landon Chun, Johnny Nguyen"
date: "2025-03-09"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```
# 1 Abstract
Our topic is questions about the NBA. We are all very interested in
the NBA, and wanted to utilize R to answer some of our questions. We decided to
look into three questions. The **first question** was "In the 2023-2024 season, who was the
biggest playoff riser". The **second question** was "At what age does performane peak in the NBA?"
and finally, the **third question** was "Does a lower defensive rating or higher offensive efficiency
correlate to a higher win percentage for top teams in the NBA in the regular season?".

We found that the highest playoff riser was surprisingly Khris Middleton We also found that
the peak performance age is 27-34. Finally, we found no correlation between defensive rating or offensive
efficiency and wins for top teams in the NBA regular season.

# 2 Introduction
As previously mentioned, our topic of analysis is the NBA. We wanted to take a deeper look
into the statistics of one of the world's most popular leagues and perhaps gain a new
perspective on different players or trends. 

This topic is interesting because it provides a unique look into the fascinating world of
professional basketball, a topic that millions of fans are interested in. The NBA pulls in
fans from across the world, so we hope that our findings will pique their interest and 
encourage them to learn more about the popular sport. Before we began our data analysis,
we knew pretty much what a basic fan knows. We knew what different statistics meant and
recognized many of the players. However, we did not know how these statistics correlate and
had many questions to answer.

In order to learn more about the NBA, we proposed three questions.

1) Which player was the biggest playoff riser in 2023-2024? 
 - We measure "playoff riser" as who had the largest increase in three distinct categories,
   P+R+A (points plus rebounds plus assists), defensive rating and offensive rating.
   (Zach Wu)
 
2) At which age does performance peak in the NBA. 
 - In order to determine the age at which
   performance peaks, we looked at offensive rating, points per game and effective field goal percentage.
   (Landon Chun)

3) Is defense overrated? Does a lower defensive rating or higher offensive efficiency rating correlate to 
a higher win percentage for top teams in the NBA in the regular season? 
- Lower defensive rating implies less points allowed in 100 possessions.
- Higher offensive efficiency means more points scored per possession.
(Johnny Nguyen)

Our intended audience is other NBA fans, likely people who are drawn to the mathematics
and statistics side of the NBA. This work is important because these fans looking at our data can 
draw new conclusions about different players, stats and more. Our data aims to increase these fan's 
interest in the NBA as a whole, by providing interesting facts and analysis.

# 3 Data source
All of our datasets came from https://www.nbastuffer.com/. This website has a bunch of stats
from 2016 to current day. To answer our questions, we used a variety of different datasets.

To answer **question 1**, we used the 2023-2024 regular season and 2023-2024 playoff datasets (both found
here: https://www.nbastuffer.com/2024-2025-nba-player-stats/). 

To answer **question 2**, we used the 2022-2023 and 2023-2024 regular season stats
(https://www.nbastuffer.com/2022-2023-nba-player-stats/ and https://www.nbastuffer.com/2023-2024-nba-player-stats/)

To answer **question 3**, we used the team and player stats in 2022-2023 and 2023-2024. 
(https://www.nbastuffer.com/2023-2024-nba-player-stats/, 
https://www.nbastuffer.com/2022-2023-nba-player-stats/, 
https://www.nbastuffer.com/2023-2024-nba-team-stats/ and 
https://www.nbastuffer.com/2022-2023-nba-team-stats/)

All of the data was collected by Serhat Ugur. The data is freely accessible to anybody
and there are no usage or access restrictions.

The datasets represent either player stats from a specified year or team stats from
a specified year. For example, in question 1 we used a dataset that had all player data
in 2024-2025.

We chose these datasets because they have a ton of well-formatted information. Each
player dataset has seemingly every player in the NBA and their stats for the overall
year. It is a very well put-together dataset. They were crucial for solving our
proposed questions as they provided us with a ton of data.


# 4 Data
For **question 1**, the regular season dataset has 657 rows and 29 columns. The playoff
dataset has 214 rows and 29 columns. In these datasets, each row represents one
player in the 2023-2024 season.

For **question 2**, the 2022-2023 dataset has 609 rows and 29 columns. The 2023-2024 dataset
has 657 rows and 29 columns. Once again, each row within these datasets represent
one player, either in the 2022-2023 or 2023-2024 season.

For **question 3**, the 2022-2023 and 2023-2024 team datasets have 30 rows and 24 columns. 
The 2022-2023 player stats dataset has 609 rows and 29 columns. The 2023-2024 dataset has 
657 rows and 29 columns. Each row within these datasets either represents a team and their stats
for a given year, or a player and their stats during a given year

Exactly which variables are relevant depends on each question, but overall the most 
relevant variables are ORtg (offensive rating), DRtg (defensive rating), P+R+A (points plus
rebounds plus assists), PpG (points per game), ApG (assists per game), OeFF (offensive effeciency)
and WIN% (win percentage).

There are some missing values in the dataset. For example, there are a few players who do not 
have offensive and defensive ratings assigned. This is likely because they are simply not
getting enough playing time in games to warrant a rating. This is fairly common, so we made
sure to filter out players like this. Additionally, there is a column called RANK that seems
to only have N/A values. This columns isn't important to our research, so was ignored. Otherwise
there were no data quality issues.

# 5 Method

### Question 1 - Who was the biggest playoff riser?
In order to clean my data, I merged the playoff and regular season datasets, with the merge key being the player name. 
I then  created 3 new columns, ORtgIncrease, DRtgIncrease and PRAIncrease. These columns 
represented the increase in offensive rating, defensive rating and P+R+A, from regular season to the
playoffs.

While processing the data, I ran into a few issues. First, some players did not have enough
playing time to get a rating, so when I was calculating their increase in offensive and defensive
rating I was getting +/- 100 (an average rating is typically around 100). In order to fix this,
I filtered out these players, as it would be an inaccurate statement to include them as "biggest playoff
risers" considering they just didn't have any data. Additionally, there were a few instances of
players names showing up multiple times (likely because they were traded mid season). In order
to fix this, I chose the first instance, as it would represent the more recent data.

To address this question, I utilized ORtg, DRtg and PRA (most important variables). I calculated the increase,
and then graphed each player in a scatter plot, oftentimes filtering out just the top 20. I created
three scatter plots using this method, before finally creating a plot with the x axis being
the increase in PRA and y axis being the decrease in defensive rating.

I did not run into any challenges other than figuring out how to plot each person in a way
that is visually appealing. Additionally, I did not use any advanced statistical approaches.

### Question 2 - At what age does performance peak
I cleaned my data by removing duplicate rows (which removes players who were traded mid season
and played for multiple teams) and filtering for players who averaged more than 5 minutes per game

I merged my datasets by player name using full join, with the merge key being each player.

Age is the most important variable for this question because I am comparing it with other stats to
determine what age range tends to be a players offensive prime.
The other stats that I used to compare against age include APG (assists per game), PPG (points per game),
ORtg (offensive rating), and eFG% (effective field goal percentage).

In order to conduct my analysis I plotted each important stat against age, and had points that
represented players.

One challenge that I faced was that I didn’t merge the columns initially so I had two of each variable
per player because there are two years worth of data. In order to solve this problem I simply created a
new column for the relevant variables by taking the mean of the desired variable from each year. For
example I created APG in my filtered dataset by taking the mean of APG.x and APG.y.

I used loess (locally estimated statistical smoothing) in my graphs which is a type of regression that
isn’t necessarily linear as it takes into account local regressions of data allowing a line that isn’t
straight and follows trends in the data.

### Question 3 - How does PER impact playoff wins
I cleaned my data by removing missing values in my datasets. I made sure to not include them in my
calculations and in doing so, I ensured that all teams had a valid average defensive rating and not just
N/A.

Because I was focusing on the top teams in the NBA, I decided to focus on the teams that made the
playoffs and therefore filtered out all the players who belong to teams that didn’t make the playoffs. 

Since in the team datasets, there isn’t a column that denotes players’ average defensive rating, I
calculated this myself by first grouping the dataset by the attribute “TEAM” and then for each team
computing the mean for the players’ defensive rating. 

When merging datasets, I made certain that the keys for both datasets can be matched with each other,
in that the team names must match each other when matching. I was surprised when in the team stats
datasets, all of the team names were listed according to their full name instead of being abbreviated
like how the player stats are. I created a separate column called “TEAMCode” and stored the
abbreviations. This allowed me to merge both the player's and team's stats 
(by.x = TEAM, by.y = TEAMCode)

For technical decisions, since I wasn’t able to find the defensive ratings for the players who don’t
have one in the dataset, I made sure to remove all missing values so that there aren’t any N/A 
in my final calculations. 

For this research question, I was interested in “TEAM”, “DRtg” (Defensive Rating), “oEFF” (Offensive
Efficiency Rating), and “WIN%” (Win Percentage Regular Season)

**Note:**

 - DRtg is a metric used to measure how many points a team allows per 100 possessions. This means that a
   lower DRtg indicates better defense. 
   
 - oEFF is a metric used to measure how effectively a team scores per 100 possessions. This metric is
   adjusted for pace and a higher oEFF rating means a more efficient offensive system. 
   
To conduct my analysis,

 - I researched what metrics are the best at identifying a team’s defensive success and
 offensive success. 
 
 - I made sure that the datasets came from a reputable and reliable source.
 
 - I made sure to understand what each number in the dataset meant. 
 
 - I made a separate column in the team datasets to include the teams’ name abbreviations so
  that I could merge the player and team stats. 

Initially, when I first merged the player and team stats based on “TEAM”, it resulted in no rows. I
identified my error in that team names in the team stats dataset were written fully and not abbreviated.
I made sure to change this by adding a separate column with all the abbreviations.


# 6 Results
### Question 1 - Who was the biggest playoff riser?
#### Loading Data
```{r}
library(tidyverse)
rs <- read_delim("NBA 2023-2024 Regular Season.csv", show_col_types = FALSE)
pl <- read_delim("NBA 2023- 2024 Playoffs.csv", show_col_types = FALSE)

md <- merge(rs, pl, by.x = "NAME", by.y = "NAME", all = FALSE, suffixes = c(".regular", ".playoffs"))
md <- md %>%
  filter(GP.playoffs >= 3 & `USG%.playoffs` >= 18.0) %>%
  distinct(NAME, .keep_all = TRUE)

md <- md %>% 
  select(NAME, `P+R+A.regular`, ORtg.regular, DRtg.regular, `P+R+A.playoffs`, ORtg.playoffs, DRtg.playoffs) %>% 
  mutate(PRAincrease = `P+R+A.playoffs` - `P+R+A.regular`, 
         ORtgIncrease = ORtg.playoffs - ORtg.regular,
         DRtgIncrease = DRtg.playoffs - DRtg.regular) %>% 
  filter(ORtg.regular != 0 & DRtg.regular != 0 & DRtgIncrease > -90)
```
#### Plot 1: X-axis = Names, Y-axis = Increase in PRA
```{r}
md %>% 
  arrange(-PRAincrease) %>% 
  head(20) %>% 
  ggplot(aes(x = reorder(NAME, PRAincrease), y = PRAincrease)) +
  geom_point() + 
  labs(x =" Player Name",
       y = "Increase in Points, Rebounds and Assists",
       title = "Players with the Highest Increase in P+R+A ")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #This line is from stack overflow because the textbook doesn't have anything about adjusting angle
  
```

This plot tells us that the player with the largest (and by a significant number) increase in 
points + rebounds + assists (PRA) from the 2023-2024 regular season to playoffs was Khris Middleton. His
PRA increased by ~12, a shockingly high value.


#### Plot 2: X-axis = Names, Y-axis = Increase in Offensive Rating
```{r}
md %>% 
  arrange(-ORtgIncrease) %>% 
  head(20) %>% 
  ggplot(aes(x = reorder(NAME, ORtgIncrease), y = ORtgIncrease)) +
  geom_point() + 
  labs(x =" Player Name",
       y = "Increase in Offensive Rating",
       title = "Players with the Highest Increase in Offensive Rating")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

This plot tells us that the player with the largest increase in offensive rating from
regular season to playoffs in 2023-2024 was Bones Hyland. In close second was Alec Burks.

#### Plot 3: X-axis = Names, Y-axis = Increase in Defensive Rating
```{r}
md %>% 
  arrange(DRtgIncrease) %>% 
  head(20) %>%
  ggplot(aes(x = reorder(NAME, DRtgIncrease), y = DRtgIncrease)) +
  geom_point() + 
  labs(x =" Player Name",
       y = "Decrease in Defensive Rating",
       title = "Players with Largest Decrease in Defensive Rating (Lower Rating = Better Defense")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

This plot shows us that the player with the largest decrease in defensive rating was Josh Okogie.

#### Plot 4: X-axis = Increase in PRA, Y-axis = Increase in Offensive Rating
```{r}
md %>% 
  ggplot(aes(x = PRAincrease, y = DRtgIncrease)) +
  geom_point() + 
  geom_text(aes(label = ifelse(NAME %in% c("Khris Middleton", "Bones Hyland", "Josh Okogie","Anthony Davis",
                                   "Luka Doncic", "Jayson Tatum"), NAME, "")), vjust = 1) + 
  labs(x = "Increase in Points + Rebounds + Assists", 
       y = "Decrease in Defensive Rating (Lower Rating = Better D)")
  
```

This plot shows us the distribution of players, where the x axis is the increase in PRA and the y axis is the
decrease in defensive rating. From this graph we can see Khris Middleton with the highest increase in PRA and 
Josh Okogie with the largest decrease in defensive rating. Some superstars' names are also 
shown for reference.

### Question 2 - At Which Age Does Performance Peak in the NBA
#### Loading Data and Manipulating Dataset
```{r}
RegStats <- read_delim("NBA 2023-2024 Regular Season.csv", show_col_types = FALSE)

RegStats2223 <- read_delim("NBA 2022-2023.csv", show_col_types = FALSE)
RegStats <- RegStats %>% distinct(NAME, .keep_all = TRUE)
RegStats2223 <- RegStats2223 %>% distinct(NAME, .keep_all = TRUE)
MergedRegSeason <-RegStats %>% 
  full_join(RegStats2223, by = "NAME")
FilteredRegStats <- MergedRegSeason %>% 
  filter(MPG.x >= 5, MPG.y >= 5, na.rm = TRUE)

FilteredRegStats <- FilteredRegStats %>% 
  mutate(AGE = (AGE.x+AGE.y)/2)
FilteredRegStats <- FilteredRegStats %>% 
  mutate(APG = (APG.x+APG.y)/2)
FilteredRegStats <- FilteredRegStats %>% 
  mutate(PPG = (PPG.x+PPG.y)/2)
FilteredRegStats <- FilteredRegStats %>% 
  mutate(`eFG%` = (`eFG%.x`+`eFG%.y`)/2)
FilteredRegStats <- FilteredRegStats %>% 
  mutate(ORtg = (ORtg.x+ORtg.y)/2)
```

#### Plot 1: X-axis = Age, Y-axis = Points Per Game
```{r}
FilteredRegStats %>% 
ggplot(aes(x = AGE, y = PPG, color = POS.x))+
  geom_point()+
  geom_smooth(method = "loess", se = FALSE, color = "blue")+
  labs(title = "Player Effectiveness Based on Age", x = "Age", y = "Points Per Game")
```

This plot shows the distribution of PPG over age. We can see that the average PPG seems
to peak around 32, and remains high from age 27-34.

#### Plot 2: X-axis = Age, Y-axis = Assists Per Game
```{r}
FilteredRegStats %>% 
ggplot(aes(x = AGE, y = APG , color = POS.x))+
  geom_point()+
  geom_smooth(method = "loess", se = FALSE, color = "blue")+
  labs(title = "Offensive Effectiveness By Age", x = "Age", y = "Assists Per Game")
```

This plot shows the how assists per game change as age increases. From this plot we see
that assists tend to consistently go up as players age.

#### Plot 3: X-axis = Age, Y-axis = Effective Field Goal Percentage
```{r}
FilteredRegStats %>% 
ggplot(aes(x = AGE, y = `eFG%`, color = POS.x))+
  geom_point()+
  geom_smooth(method = "loess", se = FALSE, color = "blue")+
  labs(title = "Offensive Effectiveness by Age", x = "Age", y = "Effective Field Goal Percentage")
```

This plot shows the average offensive effeciency as age increases. From this plot we can see
that offensive efficiency tends to peak from 24-32, and that there is a drop off on both sides of the peak.

#### Plot 4: X-axis = Age, Y-axis = Offensive Rating
```{r}
FilteredRegStats %>% 
ggplot(aes(x = AGE, y = ORtg, color = POS.x))+
  geom_point()+
  geom_smooth(method = "loess", se = FALSE, color = "blue")+
  labs(title = "Offensive Effectiveness By Age", x = "Age", y = "Offensive Rating")
```

This plot shows the average offensive rating as age increases. From this plot we can see that 
a player's offensive rating peaks from 27 to 32. Like the last plot, there are drop-offs on either side of 
the peak.

### Question 3 - How Does Offensive and Defensive Rating Impact Regular Season Wins?
#### Loading and Filtering Data
```{r}
## Load player stats for both years
playerRegular2023 <- read_delim("NBA 2023-2024 Regular Season.csv", show_col_types = FALSE)
playerRegular2022 <- read_delim("NBA 2022-2023.csv", show_col_types = FALSE)

## Load team stats for both year
teamRegular2023 <- read_delim("NBA 2023-2024 Team Regular.csv", show_col_types = FALSE)
teamRegular2022 <- read_delim("NBA 2022-2023 TEAM Regular.csv", show_col_types = FALSE)
teamPlayoff2023 <- read_delim("NBA 2023-2024 TEAM Playoffs.csv", show_col_types = FALSE)
teamPlayoff2022 <- read_delim("NBA 2022-2023 TEAM Playoffs.csv", show_col_types = FALSE)

defensiveRegular2023 <- playerRegular2023 %>% 
  filter(!is.na(DRtg)) %>% 
  filter(TEAM %in% teamPlayoff2023$TEAM) %>% 
  group_by(TEAM) %>% 
  summarise(averageDRtgRegular = mean(DRtg))

teamCode = c("Bos", "Cle", "Dal", "Den", "Ind", "Lac", "Lal", "Mia", "Mil", "Min",
             "Nor", "Nyk", "Okc", "Orl", "Phi", "Pho")

## Since the datasets are different in how team names are represented, make a new column to ensure that their both are the same when merging 
teamRegular2023Mod <- teamRegular2023 %>% 
  filter(TEAM == "Boston" | TEAM == "Cleveland" | TEAM == "Dallas" | TEAM == "Denver" | TEAM == "Indiana" | TEAM == "LA Clippers" | 
           TEAM == "LA Lakers" | TEAM == "Miami" | TEAM == "Milwaukee" | TEAM == "Minnesota" | TEAM == "New Orleans" | 
           TEAM == "New York" | TEAM == "Oklahoma City" | TEAM == "Orlando" | TEAM == "Philadelphia" | TEAM == "Phoenix") %>% 
  arrange(TEAM) %>% 
  mutate(TEAMCode = teamCode)

## Merge the two datasets together 
winVsdefRegular2023 <- merge(defensiveRegular2023, teamRegular2023Mod, by.x = "TEAM", by.y = "TEAMCode")
```

#### Plot 1: X-axis = Average Defensive Rating, Y-axis = Team Win Percentage - Regular Season 2023
Does a lower average defensive rating correlate to a higher win percentage for the regular season in 2023? (Lower defensive rating correlate to better defense)
```{r}
winVsdefRegular2023 %>% 
  ggplot(aes(x = averageDRtgRegular, y = `WIN%`)) +
    geom_point() + 
    labs(x = "Average defensive rating", 
         y = "Team win percentage", 
         title = "Does a lower average defensive rating correlate to a higher win percentage?")
  
```

This plot shows the team win percentage as defensive rating increases. From the plot 
you can see that there isn't a strong trend in one direction, if anything the win percentage
doesn't change as defensive rating increases.

#### Plot 2: X-axis = Offensive Efficiency Rating, Y-axis = Team Win Percentage - Regular Season 2023
Counter Argument: Does a higher offensive efficiency rating correlate to a higher win percentage for teams in the regular season?
```{r}
winVsdefRegular2023 %>% 
  ggplot(aes(x = oEFF, y = `WIN%`)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(x = "Offensive efficiency rating", y = "Team win percentage", title = "Does a higher offensive efficiency rating correlate a to higher win percentage?")
```

This plot shows the team win percentage as offensive efficiency rating increases. This plot shows us 
that while there is a general trend upwards as offensive efficiency rating increases, there is no
exact strong correlation. For example, the team with the highest efficiency rating has one of 
the lowest team win percentages overall.

#### Plot 3: X-axis = Average Defensive Rating, Y-axis = Team Win Percentage, Point Color = Offensive Efficiency- Regular Season 2023
Counter Argument: Does a combination of both offense and defense lead to a better win percentage? 
```{r}
winVsdefRegular2023 %>% 
  ggplot(aes(x = averageDRtgRegular, y = `WIN%`, col = oEFF)) + geom_point() + scale_color_distiller(palette = "Blues", direction = 1) + 
  labs(x = "Average defensive rating", y = "Team win percentage", col = "Offensive efficiency rating", title = "Does a combination offense and defense correspond higher win percentage?")
```

This plot shows the team win percentage on the y axis and average defensive rating on the x axis. The
color of the dot represents the offensive efficiency rating for each team. Graph shows us that there
isn't necessarily a strong trend (recall that lower defensive rating = better defense) in 2023-2024.

Calculate average defensive rating for each team in 2022
```{r}
defensiveRegular2022 <- playerRegular2022 %>% 
  filter(!is.na(DRtg)) %>% 
  filter(TEAM %in% teamPlayoff2022$TEAM) %>% 
  group_by(TEAM) %>% 
  summarise(averageDRtgRegular = mean(DRtg))
```

Merge datasets together
```{r}
## Merge the two datasets together 
winVsdefRegular2022 <- merge(defensiveRegular2022, teamRegular2022, by = "TEAM")
```

#### Plot 4: X-axis = Average Defensive Rating, Y-axis = Team Win Percentage - Regular Season 2022
Does a lower average defensive rating correlate to a higher win percentage for the regular season in 2022? (Lower defensive rating correlate to better defense)
```{r}
winVsdefRegular2022 %>% 
  ggplot(aes(x = averageDRtgRegular, y = `WIN%`)) + geom_point() + labs(x = "Average defensive rating", y = "Team win percentage", 
                                                                        title = "Does a lower average defensive rating correlate to a higher win percentage?")
  
```

This plot shows the team win percentage as average defensive rating increases in the 2022-23 season.
As defensive rating increases, it looks like there is no real correlation between rating and team win percentage.

#### Plot 5: X-axis = Offensive Efficiency Rating, Y-axis = Team Win Percentage - Regular Season 2022
Counter Argument: Does a higher offensive efficiency rating correlate to a higher win percentage for teams in the regular season?
```{r}
winVsdefRegular2022 %>% 
  ggplot(aes(x = oEFF, y = `WIN%`)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(x = "Offensive efficiency rating", y = "Team win percentage", title = "Does a higher offensive efficiency rating correlate a to higher win percentage?")
```

This plot shows the team win percentage as offensive efficiency rating increases in 2022-2023. This plot shows us 
that while there is a general trend upwards as offensive efficiency rating increases, there is no
exact strong correlation (just like in 2023-2024). For example, the team with the highest win percentage has a below
average offensive efficiency rating.

#### Plot 6: X-axis = Average Defensive Rating, Y-axis = Team Win Percentage, Point Color = Offensive Efficiency- Regular Season 2022
Counter Argument: Does a combination of both offense and defense lead to a better win percentage? 
```{r}
winVsdefRegular2022 %>% 
  ggplot(aes(x = averageDRtgRegular, y = `WIN%`, col = oEFF)) + geom_point() + scale_color_distiller(palette = "Blues", direction = 1) + 
  labs(x = "Average defensive rating", y = "Team win percentage", col = "Offensive efficiency rating", title = "Does a combination offense and defense correspond higher win percentage?")
```

This graph shows the average defensive rating vs the team win percentage, where point colors represent
offensive efficiency ratings. This graph shows that htere is a wide range of possible outcomes
and there is no distinctive trend one way or the other.

# 7 Discussion
### Question 1 - Who was the biggest playoff riser in 2023-2024
From my results, I concluded that Khris Middleton was the biggest playoff riser in the 2023-2024
NBA season. Khris Middleton clearly had the most significant increase in P+R+A. Additionally, his
defensive rating decreased (a marginal amount, but still decreased). In comparison, I could have
chosen Josh Okogie or Bones Hyland. Josh Okogie's defensive rating decreased significantly, but
unfortunately so did his offensive rating. Bones Hyland's had the highest increase in offensive
rating, but his increase in P+R+A did not reflect that. He was more efficient, but didn't do much
more than usual. Additionally, his defensive rating increased, implying he struggle defensively. 
To summarize, Khris Middleton was the biggest playoff riser because of his significant increase in 
PRA and decrease in defensive rating. Okogie and Hyland were close, but they only improved in one
aspect of their game, unlike Middleton who improved on both.

My conclusion is limited because I based "biggest playoff riser" on increase in PRA, offensive rating
and defensive rating. Realistically, these stats should not truly determine the biggest riser,
as there are many other aspects to consider (team impact, clutch time ability and more). I can
extend the results of my study fairly far, but it will fail if scrutinized heavily, since it doesn't
account for many other stats.

For future research, I would recommend calculating and using more stats for analysis. For example,
player efficiency or player true shooting percentage could be good for improving results. Overall, 
this study could just use more statistics to analyze by.

### Question 2 - At which age does performance peak in the NBA?
My findings reveal that NBA players are typically at their offensive peak between the ages of 27-34.
Although each stats have different peaks and trends, the ages of 27-34 tend to have the highest
value across the stats I analyzed. 

There are a few obvious limitations of my data such as the sample size of only two years and my 
exclusion of players who switched teams during the regular season. Due to these factors, I can
only extend my conclusion to the past couple of NBA seasons for players who were loyal to their
team in each respective season.

A direction for my future research would be to find other advanced stats to compare against age,
and also account for the minutes disparity in conuting stats such as PPG and APG.

A different direction would be to compare age against defensive effectiveness since it may be 
different than offensive effectiveness because defense tends to rely more on athleticism that 
will inevitably diminish as a player ages.

### Question 3 - How does offensive efficiency or defensive rating impact team wins in the regular season?
According to my graphs, there isn’t a significant relationship between the 3 variables: DRtg, oEFF, and WIN%.
For both 2022 and 2023, having a better defensive rating or a more efficient offense system doesn’t result in
more team wins in the regular season. Data values are spread across the graphs for both variables (DRtg and oEFF)
when isolated. When combining both together (a good defense and a good offense), there also seems to be no
relationship with team success, indicating that there must be external factors that affect team win percentage
(Team Chemistry, Coaching Strategy, Team Culture, Trades, Ownership, etc). 

Since I was only able to find 2 years' worth of players and team stats, my results are not statistically 
significant. By having more data to analyze, we can infer more about whether or not defense is a good indicator
of team success. 

For future research, I wish to incorporate more data from past years and use methods like linear regression and 
calculating the p-value to understand whether or not my results can be generalized, and where I can make an accurate
conclusion based on it. Besides this, if I were able to conduct my own research and data collection, I’d make sure to
pay close attention to things like defensive rebounding, and forced turnovers/defensive pressure because 
defensive rating doesn’t capture all of a team’s defense. I think combining multiple attributes in analyzing
a team’s defense can be valuable in seeing whether or not defense leads to more wins in the regular season. 

# 8 Summary
In conclusion, after studying recent NBA seasons in-depth, we have come to the following conclusions. The
biggest playoff riser in 2023-2024 was Khris Middleton, the peak performance age is 27-34 and finally, there
is no strong correlation between offensive efficiency or defensive rating on team wins. These conclusions will 
hopefully create a new perspective into the NBA for fans across the world, while providing interesting
takeaways, shaping the way we understand the NBA as a whole.









